Local Spectral Clustering of Density Upper Level Sets

### 2021-12-23 arxiv/appendix.tex
Updated arxiv version of paper.

\section{Fixed graph results}
\label{apdx:fixed_graph}
In this appendix, we give all results that hold with respect to an arbitrary graph $G$. For the convenience of the reader, we begin by reviewing some notation from the main text, and also introduce some new notation.

\paragraph{Notation.}
The graph $G = (V,E)$ is an undirected and connected but otherwise arbitrary graph, defined over vertices $V = \{1,\ldots,n\}$ with $m = |E|$ total edges. The adjacency matrix of $G$ is $A$, the degree matrix is $D$, and the lazy random walk matrix over $G$ is $W = (I + D^{-1}A)/2$. If the lazy random walk originates at a node $v$, the distribution of the lazy random walk $q_v^{(t)} := q(v,t;G)$ after $t$ steps is $q_v^{(t)} := e_v W^t$, with stationary distribution $\pi := \pi(G) := \lim_{t \to \infty} q_v^{(t)}$ with entries $\pi(u) = \deg(u;G)/\vol(u;G)$.

For a starting distribution $s$ (by distribution we mean a vector with non-negative entries), the PPR vector $p_s = p(s,\alpha;G)$ is the solution of
\begin{equation}
\label{eqn:ppr}
p_s = \alpha s + (1 - \alpha) p_s W.
\end{equation}
When $s = e_v$, we write $p_v := p_{e_v}$. It is easy to check that $p_s = \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t q_s^{(t)}$.  Note that $s$ need not be a probability distribution (i.e. its entries need not sum to $1$) to make sense of~\eqref{eqn:ppr}.

@@

The conductance of $V$ is abbreviated as $\Psi(G) := \Psi(V;G)$, and likewise for the local spread $s(G) := s(V;G)$. Finally, for convenience we introduce the following functionals:
\begin{equation*}
\begin{aligned}
& d_{\max}(C; G) := \max_{u \in C} \deg(u; G), && d_{\min}(C; G) := \min_{u \in C} \deg(u;G) \\
& d_{\max}(G) := d_{\max}(V;G),~~ && d_{\min}(G) := d_{\min}(V; G)
\end{aligned}
\end{equation*}
We note that $d_{\min}(G)^2 \leq d_{\min}(G) \cdot n \leq \vol(G) \leq n \cdot d_{\max}(G)$, and that for any $S \subseteq V$, $|S| \cdot d_{\min}(G) \leq \vol(S;G)$ (where $|S|$ is the cardinality of $S$.)

\paragraph{Organization.} In the following sections we establish: (Section \ref{subsec:pf_lem_zhu}) an upper bound on the misclassification error of PPR in terms of $\alpha$ and $\Phi(C;G)$ (Lemma~\ref{lem:zhu}), and an analogous result for aPPR (Corollary~\ref{cor:zhu}); (Section \ref{subsec:ppr_uniform_bounds}) a uniform bound on the perturbations of the PPR vector, to be used later in the proof of Theorem~\ref{thm:density_cluster_consistent_recovery} (consistency of PPR);  (Section \ref{subsec:lovasz_simonovits_bounds}) upper bounds on the mixedness of $q_v^{(t)}$ (as a function of $t$) and $p_v$ (as a function of $\alpha$), which will be helpful in the proofs of Proposition~\ref{prop:pointwise_mixing_time} and Theorem~\ref{thm:ppr_lb}; (Section \ref{subsec:pf_prop_pointwise_mixing_time}) an upper bound on $\tau_{\infty}(G)$ in terms of $\Psi(G)$ and $s(G)$ (Proposition~\ref{prop:pointwise_mixing_time}); and (Section \ref{subsec:ppr_spectral_partitioning}) an upper bound on the normalized cut $\Phi(\wh{C};G)$ in terms of $\Phi(C;G)$, to be used later in the proof of Theorem~\ref{thm:ppr_lb} (negative example).

\subsection{Misclassification error of clustering with PPR and aPPR}
\label{subsec:pf_lem_zhu}
For a candidate cluster $C \subseteq V$, we use the tilde-notation $\wt{G} = G[C]$ to refer to the subgraph of $G$ induced by $C$. Similarly we write $\wt{q}_v^{(t)} := q(v,t;\wt{G})$ for the $t$-step distribution of the lazy random walk over $\wt{G}$, $\wt{\pi} = \pi(G[C])$ for the stationary distribution of $\wt{q}_v^{(t)}$ (we will always assume $G[C]$ is connected), and $\wt{p}_v := p(v,\alpha;\wt{G})$ for the PPR vector over $\wt{G}$.

\paragraph{Proof of Lemma~\ref{lem:zhu}.}
As mentioned in the main text, Lemma~\ref{lem:zhu} is equivalent, up to constants, to Lemma~3.4 in \cite{zhu2013}, and the proof of Lemma~\ref{lem:zhu} proceeds along very similar lines to the proof of that lemma. In fact, we directly use the following three inequalities, derived in that work:
\begin{itemize}
	\item \textbf{(c.f. Lemma 3.2 of \cite{zhu2013})} For any seed node $v \in C$, the PPR vector is lower bounded,
	\begin{equation}
	\label{pf:zhu1}
	\wt{p}_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Corollary 3.3 of \cite{zhu2013})} For any seed node $v \in C$, there exists a so-called leakage distribution $\ell = \ell(v)$ such that $\mathrm{supp}(\ell) \subseteq C$, $\|\ell\|_1 \leq 2\Phi(C;G)/\alpha$, and 
	\begin{equation}
	\label{pf:zhu2}
	p_v(u) \geq \wt{p}_v(u) - \wt{p}_{\ell}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Lemma 3.1 of \cite{zhu2013})} There exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C;G)$ such that for any seed node $v \in C^g$, the following inequality holds
	\begin{equation}
	\label{pf:zhu3}
	p_v(C^c) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation}
\end{itemize}
We use~\eqref{pf:zhu1}-\eqref{pf:zhu3} to separately upper bound $\vol(S_{\beta,v} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$ and $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$; here $C^{\mathrm{int}} \cup C^{\mathrm{bdry}} = C$ is a partition of $C$, with
\begin{equation*}
C^{\mathrm{int}} := \Bigl\{u \in C: \deg(u;\wt{G}) > \bigl(1 - \alpha \cdot \beta \cdot \vol(C;G)\bigr) \deg(u;G) \Bigr\},
\end{equation*}
consisting of those vertices $u \in C$ with sufficient large degree in $\wt{G}$.

First we upper bound $\vol(S_{\beta,v} \setminus C;G)$. Observe that for any $u \in S_{\beta,v} \setminus C$, $p_v(u) > \beta \cdot \deg(u;G)$. Summing up over all such vertices, from~\eqref{pf:zhu3} we conclude that
\begin{equation}
\label{pf:zhu3.5}
\vol(S_{\beta,v} \setminus C; G) \leq \frac{p_v(C^c)}{\beta} \leq 2\frac{\Phi(C;G)}{\beta \cdot \alpha}.
\end{equation}

Next we upper bound $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$. From~\eqref{pf:zhu1} and~\eqref{pf:zhu2} we see that 
\begin{equation*}
p_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) - \wt{p}_{\ell}(u)~~\textrm{for all $u \in C$.}
\end{equation*}
If additionally $u \not\in S_{\beta,v}$ then $p_v(u) \leq \beta \deg(u;G)$, and for all such $u \in C \setminus S_{\beta,v}$,
\begin{equation}
\label{pf:zhu4}
\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  \beta\deg(u;G) \leq \wt{p}_{\ell}(u).
\end{equation}
On the other hand, for any $u \in C^{\mathrm{int}}$ it holds that
\begin{equation*}
\wt{\pi}(u) = \frac{\deg(u;\wt{G})}{\vol(\wt{G})} \geq \frac{\deg(u;\wt{G})}{\vol(G)} \geq \frac{(1 - \alpha \beta \vol(C;G))\deg(u;G)}{\vol(C;G)};
\end{equation*}
by plugging this in to~\eqref{pf:zhu4} we obtain
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \deg(u;G) \leq \wt{p}_{\ell}(u),~~\textrm{for all $u \in C^{\mathrm{int}} \setminus S_{\beta,v}$};
\end{equation*}
and summing over all such $u$ gives
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
\end{equation*}
The upper bounds on $\alpha$ and $\beta$ in~\eqref{eqn:zhu_condition} imply
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
\end{equation*}
and we conclude that
\begin{equation}
\label{pf:zhu5}
\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
\end{equation}

Finally, we upper bound $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$. Indeed, for any $u \in C^{\mathrm{bdry}}$,
\begin{equation*}
\frac{1}{\vol(C;G)}\sum_{w \not\in C} \1((u,w) \in E) \geq \alpha \cdot \beta \cdot \deg(u;G)
\end{equation*}
and summing over all such vertices yields
\begin{equation}
\label{pf:zhu6}
\vol(C^{\mathrm{bdry}};G) \leq \frac{1}{\alpha \beta \vol(C;G)}\sum_{\substack{u \in C^{\mathrm{bdry}} \\ w \not\in C}} \1((u,w) \in E) \leq \frac{\Phi(C;G)}{\alpha \cdot \beta}.
\end{equation} 
The claim follows upon summing the upper bounds in~\eqref{pf:zhu3.5}, \eqref{pf:zhu5} and~\eqref{pf:zhu6}.

@@

\paragraph{Proof of Corollary~\ref{cor:zhu}.}
	Recall that the upper bound~\eqref{eqn:zhu_ub} on $\vol(\wh{C} \vartriangle C;G)$ comes from combining the upper bounds on $\vol(\wh{C} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus \wh{C};G)$ and $\vol(C^{\mathrm{bdry}} \setminus \wh{C};G)$ in~\eqref{pf:zhu3.5},~\eqref{pf:zhu5} and~\eqref{pf:zhu6}. From the upper bound $p_v^{(\varepsilon)}(u) \leq p_v(u)$ for all $u \in V$, it is clear that both~\eqref{pf:zhu3.5} and~\eqref{pf:zhu6} continue to hold when the aPPR vector is used instead of the PPR vector.

	It remains only to establish an upper bound on~$\vol(C^{\mathrm{int}} \setminus \wh{C};G)$. For any $u \in C \setminus S_{\beta,v}$, from inequality~\eqref{pf:zhu2} and the lower bound $p_v^{(\varepsilon)}(u) \geq p_v(u)  - \varepsilon \deg(u;G)$ in~\eqref{eqn:appr_error} we deduce that
	\begin{equation}
	\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  (\beta + \varepsilon)\deg(u;G) \leq \wt{p}_{\ell}(u).
	\end{equation}
	Following the same steps as used in the proof of Lemma~\ref{lem:zhu} yields the following inequality:
	\begin{equation*}
	\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation*}
	The upper bounds on $\alpha,\beta$ in~\eqref{eqn:zhu_condition},  and on $\varepsilon$ in~\eqref{eqn:zhu_condition_2}, imply that
	\begin{equation*}
	\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
	\end{equation*}
	and we conclude that
	\begin{equation}
	\label{pf:zhu2_1}
	\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
	\end{equation}
	Summing the right hand sides of~\eqref{pf:zhu2},~\eqref{pf:zhu6}, and~\eqref{pf:zhu2_1} yields the claim.

\subsection{Uniform bounds on PPR}
\label{subsec:ppr_uniform_bounds}
As mentioned in our main text, in order to prove Theorem~\ref{thm:density_cluster_consistent_recovery}, we require a uniform bound on the PPR vector. Actually, we require two such bounds: for a candidate cluster $C \subseteq V$ and an alternative cluster $C' \subseteq V$, we require a lower bound on $p_v(u)$ for all $u \in C$, and an upper bound on $p_v(u')$ for all $u' \in C'$. In Lemma~\ref{lem:ppr_uniform_bound} we establish an upper bound that holds for all vertices $u$ in the interior $C_{o}$ of $C$, and a lower bound holds for all vertices $u'$ in the interior of $C_{o}'$ of $C'$; here
\begin{equation*}
C_{o} = \Bigl\{u \in C: \deg(u,\wt{G}) =  \deg(u;G)\Bigr\},~~\textrm{and}~~C_{o}'= \Bigl\{u \in C': \deg(u,G[C']) =  \deg(u;G)\Bigr\},
\end{equation*}
and we remind the reader that $\wt{G} = G[C]$. 
\begin{lemma}
	\label{lem:ppr_uniform_bound}
	Let $C$ and $C'$ be disjoint subsets of $V$, and suppose that
	\begin{equation*}
	\alpha \leq \frac{1}{2\tau_{\infty}(\wt{G})}.
	\end{equation*}
	Then there exists a set $C^g \subseteq C$ with $\vol(C^g;G) \geq \vol(C;G)/2$ such that for any $v \in C^g$,
	\begin{equation}
	\label{eqn:ppr_uniform_bound_C}
	p_v(u) \geq \frac{3}{8}\wt{\pi}(u) - \frac{2 \Phi(C;G)}{d_{\min}(\wt{G})\cdot \alpha}~~\textrm{for all $u \in C_{o}$}
	\end{equation}
	and
	\begin{equation}
	\label{eqn:ppr_uniform_bound_Cprime}
	p_v(u') \leq \frac{2\Phi(C;G)}{d_{\min}(C';G) \cdot \alpha}~~\textrm{for all $u \in C_{o}'$.}
	\end{equation}
\end{lemma}

\paragraph{``Leakage'' and ``soakage'' vectors.} To prove Lemma~\ref{lem:ppr_uniform_bound}, we will make use of the following explicit representation of the \emph{leakage} distribution $\ell$ from~\eqref{pf:zhu3}, as well as an analogously defined \emph{soakage} distribution $s$:
\begin{equation}
\label{eqn:leakage_soakage}
\begin{aligned}
\ell^{(t)} & := e_v(W \wt{I})^t(I - D^{-1}\wt{D}),~~&& \ell = \sum_{t = 0}^{\infty} (1 - \alpha)^t \ell^{(t)} \\
s^{(t)} & := e_v(W \wt{I})^t W (I - \wt{I}),~~&& s = \sum_{t = 0}^{\infty} (1 - \alpha)^t s^{(t)}.
\end{aligned}
\end{equation}
In the above, $\wt{I} \in \Reals^{n \times n}$ is a diagonal matrix with $I_{uu} = 1$ if $u \in C$ and $0$ otherwise, and $\wt{D}$ is the diagonal matrix with $\wt{D}_{uu} = \deg(u;\wt{G})$ if $u \in C$, and $0$ otherwise.

>>>>

\section{Fixed Graph Results}
\label{apdx:fixed_graph}
In this section, we give all results that hold with respect to an arbitrary graph $G$. For the convenience of the reader, we begin by reviewing some notation from the main text, and also introduce some new notation. \\

\paragraph{Notation.}
The graph $G = (V,E)$ is an undirected and connected but otherwise arbitrary graph, defined over vertices $V = \{1,\ldots,n\}$ with $m = |E|$ total edges. The adjacency matrix of $G$ is $A$, the degree matrix is $D$, and the lazy random walk matrix over $G$ is $W = (I + D^{-1}A)/2$. The lazy random walk originating at node $v \in V$ has distribution $q(v,t;G) = e_vW^t$ after $t$ steps; we use the notational shorthand $q_v^{(t)} := q(v,t;G)$. The stationary distribution of the lazy random walk is $\pi := \pi(G) := \lim_{t \to \infty} q_v^{(t)}$ is given by $\pi(u) = \deg(u;G)/\vol(u;G)$.

For a starting distribution $s$ (by distribution we mean a vector with non-negative entries), the PPR vector $p_s = p(s,\alpha;G)$ is the solution to
\begin{equation}
\label{eqn:ppr}
p_s = \alpha s + (1 - \alpha) p_s W.
\end{equation}
When $s = e_v$, we write $p_v := p_{e_v}$. It is easy to check that $p_s = \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t q_s^{(t)}$.  Note that $s$ need not be a probability distribution (i.e. its entries need not sum to $1$) to make sense of~\eqref{eqn:ppr}.

@@

The conductance of $V$ is abbreviated as $\Psi(G) := \Psi(V;G)$, and likewise for the local spread $s(G) := s(V;G)$. Finally, for convenience we introduce the following functionals:
\begin{equation*}
\begin{aligned}
& d_{\max}(C; G) := \max_{u \in C} \deg(u; G), && d_{\min}(C; G) := \min_{u \in C} \deg(u;G) \\
& d_{\max}(G) := d_{\max}(V;G),~~ && d_{\min}(G) := d_{\min}(V; G)
\end{aligned}
\end{equation*}
We note that $d_{\min}(G)^2 \leq d_{\min}(G) \cdot n \leq \vol(G) \leq d_{\max}(G) \cdot n$, and that for any $S \subseteq V$, $|S| \cdot d_{\min}(G) \leq \vol(S;G)$ (where $|S|$ is the cardinality of $S$.) \\

\emph{Organization.} In the following sections we establish: (Section \ref{subsec:pf_lem_zhu}) an upper bound on the misclassification error of PPR in terms of $\alpha$ and $\Phi(C;G)$ (Lemma~\ref{lem:zhu}), and an analogous result for aPPR (Corollary~\ref{cor:zhu}); (Section \ref{subsec:ppr_uniform_bounds}) a uniform bound on the perturbations of the PPR vector, to be used later in the proof of Theorem~\ref{thm:density_cluster_consistent_recovery} (consistency of PPR);  (Section \ref{subsec:lovasz_simonovits_bounds}) upper bounds on the mixedness of $q_v^{(t)}$ (as a function of $t$) and $p_v$ (as a function of $\alpha$), which will be helpful in the proofs of Proposition~\ref{prop:pointwise_mixing_time} and Theorem~\ref{thm:ppr_lb}; (Section \ref{subsec:pf_prop_pointwise_mixing_time}) an upper bound on $\tau_{\infty}(G)$ in terms of $\Psi(G)$ and $s(G)$ (Proposition~\ref{prop:pointwise_mixing_time}); and (Section \ref{subsec:ppr_spectral_partitioning}) an upper bound on the normalized cut $\Phi(\wh{C};G)$ in terms of $\Phi(C;G)$, to be used later in the proof of Theorem~\ref{thm:ppr_lb} (negative example).

\subsection{Misclassification Error of Clustering with PPR and aPPR}
\label{subsec:pf_lem_zhu}
For a candidate cluster $C \subseteq V$, we use the tilde-notation $\wt{G} = G[C]$ to refer to the subgraph of $G$ induced by $C$. Similarly we write $\wt{q}_v^{(t)} := q(v,t;\wt{G})$ for the $t$-step distribution of the lazy random walk over $\wt{G}$, $\wt{\pi} = \pi(G[C])$ for the stationary distribution of $\wt{q}_v^{(t)}$ (we will always assume $G[C]$ is connected), and $\wt{p}_v := p(v,\alpha;\wt{G})$ for the PPR vector over $\wt{G}$. \\

\emph{Proof of Lemma~\ref{lem:zhu}.}
As mentioned in the main text, Lemma~\ref{lem:zhu} is equivalent, up to constants, to Lemma~3.4 in \cite{zhu2013}, and the proof of Lemma~\ref{lem:zhu} proceeds along very similar lines to the proof of that lemma. In fact, we directly use the following three inequalities, derived in that work:
\begin{itemize}
	\item \textbf{(c.f. Lemma 3.2 of \cite{zhu2013})} For any seed node $v \in C$, the PPR vector is lower bounded,
	\begin{equation}
	\label{pf:zhu1}
	\wt{p}_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Corollary 3.3 of \cite{zhu2013})} For any seed node $v \in C$, there exists a so-called leakage distribution $\ell = \ell(v)$ such that $\mathrm{supp}(\ell) \subseteq C$, $\|\ell\|_1 \leq 2\Phi(C;G)/\alpha$, and 
	\begin{equation}
	\label{pf:zhu2}
	p_v(u) \geq \wt{p}_v(u) - \wt{p}_{\ell}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Lemma 3.1 of \cite{zhu2013})} There exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C;G)$ such that for any seed node $v \in C^g$, the following inequality holds
	\begin{equation}
	\label{pf:zhu3}
	p_v(C^c) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation}
\end{itemize}
We use~\eqref{pf:zhu1}-\eqref{pf:zhu3} to separately upper bound $\vol(S_{\beta,v} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$ and $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$; here $C^{\mathrm{int}} \cup C^{\mathrm{bdry}} = C$ is a partition of $C$, with
\begin{equation*}
C^{\mathrm{int}} := \Bigl\{u \in C: \deg(u;\wt{G}) > \bigl(1 - \alpha \cdot \beta \cdot \vol(C;G)\bigr) \deg(u;G) \Bigr\},
\end{equation*}
consisting of those vertices $u \in C$ with sufficient large degree in $\wt{G}$.

First we upper bound $\vol(S_{\beta,v} \setminus C;G)$. Observe that for any $u \in S_{\beta,v} \setminus C$, $p_v(u) > \beta \cdot \deg(u;G)$. Summing up over all such vertices, from~\eqref{pf:zhu3} we conclude that
\begin{equation}
\label{pf:zhu3.5}
\vol(S_{\beta,v} \setminus C; G) \leq \frac{p_v(C^c)}{\beta} \leq 2\frac{\Phi(C;G)}{\beta \cdot \alpha}.
\end{equation} 
Next we upper bound $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$. From~\eqref{pf:zhu1} and~\eqref{pf:zhu2} we see that 
\begin{equation*}
p_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) - \wt{p}_{\ell}(u)~~\textrm{for all $u \in C$.}
\end{equation*}
If additionally $u \not\in S_{\beta,v}$ then $p_v(u) \leq \beta \deg(u;G)$, and for all such $u \in C \setminus S_{\beta,v}$,
\begin{equation}
\label{pf:zhu4}
\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  \beta\deg(u;G) \leq \wt{p}_{\ell}(u).
\end{equation}
On the other hand, for any $u \in C^{\mathrm{int}}$ it holds that
\begin{equation*}
\wt{\pi}(u) = \frac{\deg(u;\wt{G})}{\vol(\wt{G})} \geq \frac{\deg(u;\wt{G})}{\vol(G)} \geq \frac{(1 - \alpha \beta \vol(C;G))\deg(u;G)}{\vol(C;G)};
\end{equation*}
by plugging this in to~\eqref{pf:zhu4} we obtain
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \deg(u;G) \leq \wt{p}_{\ell}(u),~~\textrm{for all $u \in C^{\mathrm{int}} \setminus S_{\beta,v}$};
\end{equation*}
and summing over all such $u$ gives
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
\end{equation*}
The upper bounds on $\alpha$ and $\beta$ in~\eqref{eqn:zhu_condition} imply
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
\end{equation*}
and we conclude that
\begin{equation}
\label{pf:zhu5}
\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
\end{equation}
Finally, we upper bound $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$. Indeed, for any $u \in C^{\mathrm{bdry}}$,
\begin{equation*}
\frac{1}{\vol(C;G)}\sum_{w \not\in C} \1((u,w) \in E) \geq \alpha \cdot \beta \cdot \deg(u;G)
\end{equation*}
and summing over all such vertices yields
\begin{equation}
\label{pf:zhu6}
\vol(C^{\mathrm{bdry}};G) \leq \frac{1}{\alpha \beta \vol(C;G)}\sum_{\substack{u \in C^{\mathrm{bdry}} \\ w \not\in C}} \1((u,w) \in E) \leq \frac{\Phi(C;G)}{\alpha \cdot \beta}.
\end{equation} 
The claim follows upon summing the upper bounds in~\eqref{pf:zhu3.5}, \eqref{pf:zhu5} and~\eqref{pf:zhu6}. \qed

@@

\noindent \emph{Proof of Corollary~\ref{cor:zhu}.}
Recall that the upper bound~\eqref{eqn:zhu_ub} on $\vol(\wh{C} \vartriangle C;G)$ comes from combining the upper bounds on $\vol(\wh{C} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus \wh{C};G)$ and $\vol(C^{\mathrm{bdry}} \setminus \wh{C};G)$ in~\eqref{pf:zhu3.5},~\eqref{pf:zhu5} and~\eqref{pf:zhu6}. From the upper bound $p_v^{(\varepsilon)}(u) \leq p_v(u)$ for all $u \in V$, it is clear that both~\eqref{pf:zhu3.5} and~\eqref{pf:zhu6} continue to hold when the aPPR vector is used instead of the PPR vector.

It remains only to establish an upper bound on~$\vol(C^{\mathrm{int}} \setminus \wh{C};G)$. For any $u \in C \setminus S_{\beta,v}$, from inequality~\eqref{pf:zhu2} and the lower bound $p_v^{(\varepsilon)}(u) \geq p_v(u)  - \varepsilon \deg(u;G)$ in~\eqref{eqn:appr_error} we deduce that
\begin{equation}
\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  (\beta + \varepsilon)\deg(u;G) \leq \wt{p}_{\ell}(u).
\end{equation}
Following the same steps as used in the proof of Lemma~\ref{lem:zhu} yields the following inequality:
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
\end{equation*}
The upper bounds on $\alpha,\beta$ in~\eqref{eqn:zhu_condition},  and on $\varepsilon$ in~\eqref{eqn:zhu_condition_2}, imply that
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
\end{equation*}
and we conclude that
\begin{equation}
\label{pf:zhu2_1}
\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
\end{equation}
Summing the right hand sides of~\eqref{pf:zhu2},~\eqref{pf:zhu6}, and~\eqref{pf:zhu2_1} yields the claim. \qed

\subsection{Uniform Bounds on PPR}
\label{subsec:ppr_uniform_bounds}
As mentioned in our main text, in order to prove Theorem~\ref{thm:density_cluster_consistent_recovery}, we require a uniform bound on the PPR vector. Actually, we require two such bounds: for a candidate cluster $C \subseteq V$ and an alternative cluster $C' \subseteq V$, we require a lower bound on $p_v(u)$ for all $u \in C$, and an upper bound on $p_v(u')$ for all $u' \in C'$. In Lemma~\ref{lem:ppr_uniform_bound} we establish an upper bound that holds for all vertices $u$ in the interior $C_{o}$ of $C$, and a lower bound holds for all vertices $u'$ in the interior of $C_{o}'$ of $C'$; here
\begin{equation*}
C_{o} = \Bigl\{u \in C: \deg(u,\wt{G}) =  \deg(u;G)\Bigr\},~~\textrm{and}~~C_{o}'= \Bigl\{u \in C': \deg(u,G[C']) =  \deg(u;G)\Bigr\},
\end{equation*}
and we remind the reader that $\wt{G} = G[C]$. 
\begin{lemma}
	\label{lem:ppr_uniform_bound}
	Let $C$ and $C'$ be disjoint subsets of $V$, and suppose that
	\begin{equation*}
	\alpha \leq \frac{1}{2\tau_{\infty}(\wt{G})}.
	\end{equation*}
	Then there exists a set $C^g \subseteq C$ with $\vol(C^g;G) \geq \vol(C;G)/2$ such that for any $v \in C^g$,
	\begin{equation}
	\label{eqn:ppr_uniform_bound_C}
	p_v(u) \geq \frac{3}{8}\wt{\pi}(u) - \frac{2 \Phi(C;G)}{d_{\min}(\wt{G})\cdot \alpha}~~\textrm{for all $u \in C_{o}$}
	\end{equation}
	and
	\begin{equation}
	\label{eqn:ppr_uniform_bound_Cprime}
	p_v(u') \leq \frac{2\Phi(C;G)}{d_{\min}(C';G) \cdot \alpha}~~\textrm{for all $u \in C_{o}'$.}
	\end{equation}
\end{lemma}

\noindent\emph{``Leakage'' and ``soakage'' vectors.} To prove Lemma~\ref{lem:ppr_uniform_bound}, we will make use of the following explicit representation of the \emph{leakage} distribution $\ell$ from~\eqref{pf:zhu3}, as well as an analogously defined \emph{soakage} distribution $s$:
\begin{equation}
\label{eqn:leakage_soakage}
\begin{aligned}
\ell^{(t)} & := e_v(W \wt{I})^t(I - D^{-1}\wt{D}),~~&& \ell = \sum_{t = 0}^{\infty} (1 - \alpha)^t \ell^{(t)} \\
s^{(t)} & := e_v(W \wt{I})^t W (I - \wt{I}),~~&& s = \sum_{t = 0}^{\infty} (1 - \alpha)^t s^{(t)}.
\end{aligned}
\end{equation}
In the above, $\wt{I} \in \Reals^{n \times n}$ is a diagonal matrix with $I_{uu} = 1$ if $u \in C$ and $0$ otherwise, and $\wt{D}$ is the diagonal matrix with $\wt{D}_{uu} = \deg(u;\wt{G})$ if $u \in C$, and $0$ otherwise.

These quantities admit a natural interpretation in terms of random walks. For $u \in C$, $\ell^{(t)}(u)$ is the probability that a lazy random walk over $G$ originating at $v$ stays within $C$ for $t$ steps, arriving at $u$ on the $t$th step, and then ``leaks out'' of $C$ on the $(t + 1)$st step. On the other hand, for $u \not\in C$, $s^{(t)}(u)$ is the probability that a lazy random walk over $G$ originating at $v$ stays within $C$ for $t$ steps and is then ``soaked up'' into $u$ on the $(t + 1)$st step. The vectors $\ell$ and $s$ then give the total mass leaked and soaked, respectively, by the PPR vector.
